---
title: Bloomberg
author: Olivier Bauthéac
date: ""
output: rmarkdown::html_vignette
bibliography: literature_files/references.bib
vignette: >
  %\VignetteIndexEntry{Bloomberg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<style> body {text-align: justify} </style>

```{r, setup, echo = F, message = F, warning = F}
knitr::opts_chunk$set(collapse = T, comment = "#>", warning = F, message = F, eval = T)
folder <- "literature_files"; dir.create(folder)
download.file("https://www.dropbox.com/s/htnd7o9nnkk8ng8/references.bib?dl=1", paste(folder, "references.bib", sep = "/"))
path <- here::here("development", "storethat.sqlite")
```


The [finRes](https://bautheac.github.io/finRes/) suite is organised around the data-science pipeline where preprocessing, including data collection and wrangling, plays a major role. [finRes](https://bautheac.github.io/finRes/) addresses the issue in two complementary packages that work in conjunction with most of the [finRes](https://bautheac.github.io/finRes/) dataset packages.  
On the one hand the [pullit](https://bautheac.github.io/pullit/) package provides tools for data collection from Bloomberg. It returns clean and tidy, ready-to-use, data objects for other packages further down the pipeline to work with. On the other hand the [storethat](https://bautheac.github.io/storethat/) package works in concert with [fewISOs](https://bautheac.github.io/fewISOs/) and [GICS](https://bautheac.github.io/GICS/) to help store the data retrieved for off-Bloomberg consumption in R.  
Both [pullit](https://bautheac.github.io/pullit/) and [storethat](https://bautheac.github.io/storethat/) work in tandem with the [BBGsymbols](https://bautheac.github.io/BBGsymbols/) package. The latter plays a central role in [finRes](https://bautheac.github.io/finRes/) where it helps [pullit](https://bautheac.github.io/pullit/) in interacting with Bloomberg via the interface provided by the [Rblpapi](https://github.com/Rblp/Rblpapi) package [@Armstrong_Rblpapi] and [storethat](https://bautheac.github.io/storethat/) in storing the data retrieved.


## pullit
[pullit](https://bautheac.github.io/pullit/) together with [storethat](https://bautheac.github.io/storethat/) are the two workhorses for data ETL (extract, transform, load) in the [finRes](https://bautheac.github.io/finRes/) suite.
Using carefully selected Bloomberg datafields from [BBGsymbols](https://bautheac.github.io/BBGsymbols/) in tandem with @Armstrong_Rblpapi's Bloomberg interface [pullit](https://bautheac.github.io/pullit/) provides the R user with easy access to Bloomberg financial data for a number of financial instruments including, at the time of writing, equity and equity-like securities, funds with the category encompassing any money-managing entity, indexes, as well as futures series & term structure individual contracts.  

For each instrument category, [pullit](https://bautheac.github.io/pullit/) allows data retrieval for a corresponding category specific set of data types. An active Bloomberg connection is required to retrieve Bloomberg financial data using [pullit](https://bautheac.github.io/pullit/).

```{r `globals bbg`, warnings = F, message = F, eval = F, echo = F}
library(pullit); library(lubridate)
# end <- Sys.Date(); start <- end - years(2L)
start <- "2016-01-01"; end <- "2017-12-31"
```

```{r `globals storethat`, warnings = F, message = F, eval = T}
library(pullit); library(lubridate)

start <- "2016-01-01"; end <- "2017-12-31"
```

### equity
Equity data comes in three major categories in [pullit](https://bautheac.github.io/pullit/), market, book and info. For a given corporation market data records stock market activity while book data records business activity in the form of financial statements commonly referred to as 'books' and info refers to qualitative information. See the `fields` dataset in [BBGsymbols](https://bautheac.github.io/BBGsymbols/) for a detailed list of the Bloomberg datafields available.

#### market
Retrieve market data from Bloomberg for the BP (BP/ LN Equity), Weir group (WEIR LN Equity) and Apple (AAPL US Equity) corporations with:
```{r `equity market bbg`, eval = F, echo = T}
equity_tickers <- c('BP/ LN Equity', 'WEIR LN Equity', 'AAPL US Equity')

equity_market <- pull_equity_market(source = "Bloomberg", equity_tickers, start, end, verbose = F)
equity_market
```

```{r `equity market storethat`, eval = T, echo = F}
equity_tickers <- c('BP/ LN Equity', 'WEIR LN Equity', 'AAPL US Equity')

equity_market <- pull_equity_market(source = "storethat", equity_tickers, start, end, verbose = F, file = path)
equity_market
```

#### books
For financial statements data [BBGsymbols](https://bautheac.github.io/BBGsymbols/) replicates the Bloomberg 'financial analysis' monitor (FA \<GO\>). At the time of writing, this includes balance sheet, cash flow statement, income statement as well as 'key stats' that gathers broad summary figures and 'ratios' that includes popular financial ratios.

Retrieve the corresponding data for the abovementioned corporations with:  

##### __balance sheet__
```{r `equity balance sheet bbg`, eval = F, echo = T}
equity_BS <- pull_equity_book(source = "Bloomberg", book = "balance sheet", equity_tickers, 
                              start, end, verbose = F)
equity_BS
```

```{r `equity balance sheet storethat`, eval = T, echo = F}
equity_BS <- pull_equity_book(source = "storethat", book = "balance sheet", equity_tickers, 
                              start, end, verbose = F, file = path)
equity_BS
```

##### __cash flow statement__
```{r `equity cash flow statement bbg`, eval = F, echo = T}
equity_CF <- pull_equity_book(source = "Bloomberg", book = "cash flow statement", equity_tickers, 
                              start, end, verbose = F)
equity_CF
```
```{r `equity cash flow statement storethat`, eval = T, echo = F}
equity_CF <- pull_equity_book(source = "storethat", book = "cash flow statement", equity_tickers, 
                              start, end, verbose = F, file = path)
equity_CF
```
##### __income statement__
```{r `equity income statement bbg`, eval = F, echo = T}
equity_IS <- pull_equity_book(source = "Bloomberg", book = "income statement", equity_tickers, 
                              start, end, verbose = F)
equity_IS
```
```{r `equity income statement storethat`, eval = T, echo = F}
equity_IS <- pull_equity_book(source = "storethat", book = "income statement", equity_tickers, 
                              start, end, verbose = F, file = path)
equity_IS
```
##### __key stats__
```{r `equity key stats bbg`, eval = F, echo = T}
equity_KS <- pull_equity_book(source = "Bloomberg", book = "key stats", equity_tickers, 
                              start, end, verbose = F)
equity_KS
```
```{r `equity key stats storethat`, eval = T, echo = F}
equity_KS <- pull_equity_book(source = "storethat", book = "key stats", equity_tickers, 
                              start, end, verbose = F, file = path)
equity_KS
```
##### __ratios__
```{r `equity ratios bbg`, eval = F, echo = T}
equity_R <- pull_equity_book(source = "Bloomberg", book = "ratios", equity_tickers, 
                             start, end, verbose = F)
equity_R
```
```{r `equity ratios storethat`, eval = T, echo = F}
equity_R <- pull_equity_book(source = "storethat", book = "ratios", equity_tickers, 
                             start, end, verbose = F, file = path)
equity_R
```

#### info
'info' encompasses a range of contemporaneous qualitative information including, but not limited to, firm's name, security type, exchange where the security trades, country of incorporation, etc.

Retrieve the corresponding data for the abovementioned corporations with:
```{r `equity info bbg`, eval = F, echo = T}
equity_info <- pull_equity_info(source = "Bloomberg", equity_tickers, verbose = F)
equity_info
```
```{r `equity info storethat`, eval = T, echo = F}
equity_info <- pull_equity_info(source = "storethat", equity_tickers, verbose = F, file = path)
equity_info
```

### fund
Fund data only comes in two categories, market and info. For a given fund market data records stock market activity while info data records contemporaneous qualitative information that includes a wide range of fund characteristics. Market historical data as well as contemporaneous qualitative data not only allow for an in depth assessment of the fund performance but potentially also the mapping of the performance to fund characteristics. See the `fields` dataset in [BBGsymbols](https://bautheac.github.io/BBGsymbols/) for a detailed list of the Bloomberg datafields available.  

#### market
Retrieve market data from Bloomberg for the SPDR S&P 500 ETF Trust (SPY US Equity):
```{r `fund market bbg`, eval = F, echo = T}
fund_tickers <- "SPY US Equity"

fund_market <- pull_fund_market(source = "Bloomberg", fund_tickers, start, end, verbose = F)
fund_market
```

```{r `fund market storethat`, eval = T, echo = F}
fund_tickers <- "SPY US Equity"

fund_market <- pull_fund_market(source = "storethat", tickers = fund_tickers, 
                                start = start, end = end, verbose = F, file = path)
fund_market
```

#### info
Retrieve the corresponding qualitative data for the abovementioned funds with:
```{r `fund info bbg`, eval = F, echo = T}
fund_info <- pull_fund_info(source = "Bloomberg", fund_tickers, verbose = F)
fund_info
```
```{r `fund info storethat`, eval = T, echo = F}
fund_info <- pull_fund_info(source = "storethat", fund_tickers, verbose = F, file = path)
fund_info
```

### futures
Futures data comes in three major categories in [pullit](https://bautheac.github.io/pullit/), market, CFTC and info. For a given futures series market data records futures market price activity while CFTC data records market positions. 'info' on the other hand records contemporaneous qualitative information for the corresponding futures series, including but not limited to series name, term structure length, contract size, contract unit, etc. See the `fields` dataset in [BBGsymbols](https://bautheac.github.io/BBGsymbols/) for a detailed list of the Bloomberg datafields available.

#### market
[pullit](https://bautheac.github.io/pullit/) helps to retrieve both term structure as well as aggregated market data for futures. Term structure data records market activity for individual futures term structure contracts while aggregated data records market activity measures that are aggregated over the whole term structure for a particular futures series.

##### __term structure__
Retrieve market data from Bloomberg for the five first term structure contracts on the Corn - #2-yellow (C A Comdty) and silver (SIA Comdty) futures series where the futures chains are constructed by rolling on liquidity (with active contract: "A") with no roll adjustment (none: "N") using:
```{r `futures term structure bbg`, echo = T, eval = F}
futures_tickers <- c("C A Comdty", "SIA Comdty")

futures_TS <- pull_futures_market(source = "Bloomberg", type = "term structure", 
                                  active_contract_tickers = futures_tickers, 
                                  start, end, TS_positions = 1L:5L, roll_type = "A", roll_days = 0L, 
                                  roll_months = 0L, roll_adjustment = "N", verbose = F)
futures_TS
```

```{r `futures term structure storethat`, echo = F, eval = T}
futures_tickers <- c("C A Comdty", "SIA Comdty")

futures_TS <- pull_futures_market(source = "storethat", type = "term structure", 
                                  active_contract_tickers = futures_tickers, 
                                  start, end, TS_positions = 1L:5L, roll_type = "A", roll_days = 0L, 
                                  roll_months = 0L, roll_adjustment = "N", verbose = F, file = path)
futures_TS
```

##### __aggregate__
Retrieve the corresponding aggregated futures market data with:
```{r `futures aggregated bbg`, echo = T, eval = F}
futures_agg <- pull_futures_market(source = "Bloomberg", type = "aggregate", 
                                   active_contract_tickers = futures_tickers, start, end, verbose = F)
futures_agg
```
```{r `futures aggregated storethat`, echo = F, eval = T}
futures_agg <- pull_futures_market(source = "storethat", type = "aggregate", 
                                   active_contract_tickers = futures_tickers, start, end, verbose = F, file = path)
futures_agg
```

#### CFTC
[pullit](https://bautheac.github.io/pullit/) helps to retrieve CFTC futures market position data from Bloomberg. The Commodity Futures Trading Commission (CFTC) publishes the Commitments of Traders (COT) reports to help the public understand market dynamics. Specifically, the COT reports provide a breakdown of each Tuesday's open interest for futures and options on futures markets in which 20 or more traders hold positions equal to or above the reporting levels established by the CFTC. See the `fields` and `tickers_CFTC` datasets in [BBGsymbols](https://bautheac.github.io/BBGsymbols/) for details.

Retrieve the corresponding futures market position data with:
```{r `futures CFTC bbg`, echo = T, eval = F}
futures_CFTC <- pull_futures_CFTC(source = "Bloomberg", active_contract_tickers = futures_tickers, 
                                  start, end, verbose = F)
futures_CFTC
```
```{r `futures CFTC storethat`, echo = F, eval = T}
futures_CFTC <- pull_futures_CFTC(source = "storethat", active_contract_tickers = futures_tickers, 
                                  start, end, verbose = F, file = path)
futures_CFTC
```

#### info
'info' encompasses a range of contemporaneous qualitative information on the underlying futures series including, but not limited to, name for the series' underlying, trading exchange, term structure length, contract size, etc.

Retrieve the corresponding data for the abovementioned futures series with:
```{r `futures info bbg`, echo = T, eval = F}
futures_info <- pull_futures_info(source = "Bloomberg", futures_tickers, verbose = F)
futures_info
```
```{r `futures info storethat`, echo = F, eval = T}
futures_info <- pull_futures_info(source = "storethat", futures_tickers, verbose = F, file = path)
futures_info
```


## accessors
The functions above return objects that not only carry the retrieved financial data but also complementary information that can be accessed using bespoke accessor methods. Historical data functions for example return objects that carry a 'tickers' dataframe that indicates the tickers for which some data have been found, a 'fields' dataframe that indicates the data fields for which data has been found, a 'data' dataframe that hosts the retrieved data as well as a character vector hosting the original call to the function. A `get_periods()` method complements the accessor methods by indicating the start and end dates between which data have been found for each ticker and data field:

```{r `futures show`}
futures_TS
```

Access each slot using the appropriate accessor:
```{r `futures get_tickers`}
tickers <- get_active_contract_tickers(futures_TS)
tickers
```

```{r `futures get_fields`}
fields <- get_fields(futures_TS)
head(fields)
```

```{r `futures get_data`}
data <- pullit::get_data(futures_TS)
data
```

```{r `futures get_call`}
call <- pullit::get_call(futures_TS)
call
```


## storethat

All the objects above can be stored in a bespoke database for later off-Bloomberg consumption. The [storethat](https://bautheac.github.io/storethat/) package makes the process seamless by providing befitted storing methods:

```{r `storethat store`, eval = F}
library(storethat)

db_create()

path = "~/storethat.sqlite"

db_store(object = futures_TS, file = path, verbose = F)
db_store(object = fund_market, file = path, verbose = F)
```

Every function above has an equivalent for retrieving data from a [storethat](https://bautheac.github.io/storethat/) database that can be accessed by swapping the 'BBG' prefix for 'storethat'; i.e. 'pull_futures_market' becomes 'pull_futures_market'. Function parameters are identical for both data sources and the returned objects have alike characteristics:

```{r `storethat retrieve`}
equity_market <- pull_equity_market(source = "storethat", equity_tickers, start, 
                                    end, verbose = F, file = path)
equity_market
```

Updating a [storethat](https://bautheac.github.io/storethat/) database is equally straightforward with [pullit](https://bautheac.github.io/pullit/). Update the equity content of the database with: 
```{r `storethat update all`, eval = F}
storethat_update(instrument = "equity", file = path)
```
The function pulls data from Bloomberg for all the equity tickers and fields already present in the [storethat](https://bautheac.github.io/storethat/) database and updates the corresponding series up to the current system date.

Narrower updates are also allowed:
```{r `storethat update some`, eval = F}
storethat_update(instrument = "equity", book = "market")
```


## plotit
The [plotit](https://bautheac.github.io/plotit/) package, also part of the [finRes](https://bautheac.github.io/finRes/) suite, provides plot methods for some [pullit](https://bautheac.github.io/pullit/) data objects including, at the time of writing, futures term structure (FuturesTS) and fund market (FundMarket) objects.

### futures term structure
Plot a futures series term structure dynamics with:
```{r `plot term structure`, fig.width = 7.5, fig.height = 5.5, fig.fullwidth = T}
library(plotit)

plot(object = futures_TS, ticker = "C A Comdty")
```

### fund market
Plot historical fund performance with:
```{r `plot performance`, fig.width = 7.5, fig.height = 6.5, fig.fullwidth = T}
plot(object = fund_market, ticker = "SPY US Equity")
```
  
  
  
  
## references
